
---
title: "Homework 3"
author: "Beibei Cao"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include = FALSE}
library(tidyverse)
library(arsenal)
library(readr)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%",
	warning = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```

## Problem 2

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF).

Load, tidy, and otherwise wrangle the accelerometer data.
```{r message = FALSE}
# load the accel csv, transform the df from wide to long format
accel_df = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
   cols = starts_with("activity_"),
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>% 
  # encode data with reasonable variable classes, make the weekday ordered, add the weekend column
  mutate(
    week = as.integer(week),
    day_id = as.integer(day_id),
    day = as.factor(day),
    day = ordered(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")),
    minute = as.integer(minute),
    weekend = if_else(day %in% c("Saturday", "Sunday"), TRUE, FALSE)
  ) %>% 
  # sort by week, day and minute
  arrange(week, day, minute)
```

```{r}
# preview the df
accel_df
```


The `accel_df` contains ``r nrow(accel_df)`` observations of ``r ncol(accel_df)`` variables related to accelerometer data collected on the male. The `week` is an integer variable indicating the nth week of the observation  (``r min(accel_df$week)` - `r max(accel_df$week)``); the `day_id` is an integer variable indicating a unique id of the specific day from ``r min(accel_df$day_id)`` to ``r max(accel_df$day_id)``; the `day` is an ordered factor variable indicate the day of the week; the `activity_count` is a double numeric variable represents the activity counts in each minute ranging from ``r min(accel_df$activity_count)`` to `r max(accel_df$activity_count)`; the `weekend` is a logical variable indicating whether the day is a weekend or not.

Using the tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. 

```{r message = FALSE}
# group by week and day
activity_tbl = 
  accel_df %>% 
  group_by(week, day) %>% 
  # add up activity count of each day
  summarise(day_sum = sum(activity_count)) %>% 
  # format the table into wide 
  pivot_wider(
    names_from = day,
    values_from = day_sum
  ) %>% 
  knitr::kable()
```

```{r}
# show table
activity_tbl  
```

Visualize the trends of the total activity in the five week on each day.

```{r message = FALSE}
theme_set(theme_minimal() + theme(legend.position = "right"))
# visualize the trends of the total activity variable
 accel_df %>% 
  group_by(week, day) %>% 
  # add up activity count of each day
  summarise(day_sum = sum(activity_count)) %>% 
  ggplot(aes(x = day ,y = day_sum, group = week, color = week)) + 
  geom_line() + 
  geom_point() +
  labs(
    title = "Total Activity Counts of Each Day in the Week",
    x = "Day of the Week",
    y = "Day Activity Counts",
    color = "Day of the week",
    caption = "Accelerometer data collected from a 63 year-old male with BMI 25."
    ) 
  
```

No significant trend could be observed except for that in week 4 and week 5 very little activity was observed on Saturday and not many activities on Sunday. 

Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. 

```{r message = FALSE}
accel_df %>% 
  group_by(day, minute) %>% 
  # add up activity count of each day in the week across five weeks
  summarise(weekday_sum = sum(activity_count)) %>% 
  # plot with activity count of each day across five weeks against day time
  ggplot(aes(x = minute, y = weekday_sum, color = day)) + 
  geom_point(alpha = 0.3, size = 0.5) +
  geom_smooth(size = 0.5, se = FALSE) +
  labs(
    title = "24-Hour Activity Time Courses for Each Day",
    x = "24-hour of the Day (h)",
    y = "Activity Level",
    color = "Day of a Week",
    caption = "Accelerometer data collected from a 63 year-old male with BMI 25."
    ) +
  scale_x_continuous(
    breaks = c(seq(0, 1440, by = 60)),
    labels = c(seq(0, 24, by = 1))
    ) 

```

Generally, low activities levels could be observed from 23 pm to 5 am, when it was probably that the man was asleep. The activity levels peaked a little bit around the middle of each day (especially Sunday) and the end of each day (especially Friday).
















